---
title: "Peer Assessment 1"
author: "mcosmr"
date: "Tuesday, February 10, 2015"
output: html_document
---
#Peer Assessment 1
## PART 1
```{r}
#Read and preprocess data 
activity_data <- read.csv("activity.csv", sep = ",")

##Change class of Date Colum (Factor) to class "Date"
activity_data$date <- as.Date(activity_data$date, format="%Y-%m-%d")

```
## PART 2
```{r}

## Calculate the total number of steps taken per day (ignoring the missing values in the dataset)
Total_steps <- aggregate(steps ~ date, data = activity_data, na.rm =TRUE, sum)
```

###histogram of the total number of steps taken each day
`
```{r hist1, fig.keep= "all", fig= TRUE, fig.show= "asis", fig.path = "figure/", include=TRUE,echo = TRUE}
hist(Total_steps$steps, xlab = "Total Steps", main = "Steps each day")

```
###Calculate and report the mean and median of the total number of steps taken per day
  
```{r, echo = TRUE}

median_steps_day <- median(Total_steps$steps)
mean_steps_day <- mean(Total_steps$steps)

median_steps_day
mean_steps_day
```
## PART3
####Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days 

```{r plot, fig.keep= "all", fig= TRUE, fig.show= "asis", fig.path = "figure/", include=TRUE, echo = TRUE}
##Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) 
##and the average number of steps taken, averaged across all days (y-axis)

time_series <- aggregate(steps ~ interval, data = activity_data, na.rm =TRUE, mean)

plot(time_series$interval, time_series$steps, type = "l", xlab = "Interval", ylab = "Average Steps")

```
### 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

```{r, echo = TRUE}
#Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
time_series[time_series$steps == max(time_series$steps),]


```

##PART 4
```{r, echo = TRUE}
#Calculate and report the total number of missing values in the dataset

NA_count <- sum(is.na(activity_data))
NA_count
```

```{r, echo = TRUE}
#filling in all of the missing values in the dataset. 
# For example, you could use the mean/median for that day,
#or the mean for that 5-minute interval, etc.
#index_Na <- which(is.na(activity_data))

#first, replacing NA values in data steps with mean for that interval
replace_NA<- replace(activity_data[, "steps"], is.na(activity_data[, "steps"])&activity_data$interval %in% time_series$interval , time_series[,"steps"])
#create a new data set with no NA values

data2 <- activity_data
data2$steps <- replace_NA
```

```{r histog, fig.keep= "all", fig= TRUE, fig.show= "asis", fig.path = "figure/", include=TRUE, echo = TRUE}
#Make a histogram of the total number of steps taken each day 
#and Calculate and report the mean and median total number of steps taken per day.

Total_steps2 <- aggregate(steps ~ date, data = data2, sum)

hist(Total_steps2$steps, xlab = "Total Steps", main = "Steps each day")

median2_steps_day <- median(Total_steps2$steps)
mean2_steps_day <- mean(Total_steps2$steps)
```

```{r, echo = TRUE}
#comparing both data

par(mfrow = c(2,1))
```

```{r histogra, fig.keep= "all", fig= TRUE, fig.show= "asis", fig.path = "figure/", include=TRUE, echo=TRUE}

hist(Total_steps$steps, xlab = "Total Steps", main ="NA removed")
```

```{r histogram, fig.keep= "all", fig= TRUE, fig.show= "asis", fig.path = "figure/", include=TRUE, echo=TRUE}
hist(Total_steps2$steps, xlab = "Total Steps ", main = "NA Filled in")


## 5-minute interval, on average across all the days in the dataset, maximum number of steps

time_series2 <- aggregate(steps ~ interval, data = data2, mean)
time_series2[time_series2$steps == max(time_series2$steps),]


```
## PART 5
```{r, echo = TRUE}

#Create a new factor variable in the dataset with two levels - "weekday" and "weekend" 
#indicating whether a given date is a weekday or weekend day.
Sys.setlocale("LC_TIME", "English")
data2$week_days<- weekdays(data2$date)
data2$week_days <- ifelse(data2$week_days == "Saturday" | data2$week_days == "Sunday", "Weekend", "Weekday")
data2$week_days <- as.factor(data2$week_days)


```
```{r, echo = TRUE}
##Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) ##and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

serie_weekday<- aggregate(data2$steps, list(data2$interval, data2$week_days), mean)
colnames(serie_weekday) <- c("interval", "week_day", "steps")

```
###panel plot
```{r qplot, fig.keep= "all", fig= TRUE, fig.show= "asis", fig.path = "figure/", include=TRUE, echo = TRUE}

library(ggplot2)
qplot(interval, steps, data = serie_weekday, geom = "line", facets = week_day~.)

```
